---
title: "Worksheet-4a in R"
author: "Trexan Sanico"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. The table below shows the data about shoe size and height. Create a data frame.

```{r}
 shoe_size <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)
  height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0)
  gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
  
  data <- data.frame(Shoe_size = shoe_size, Height = height, Gender = gender)

```
#a. Describe the data.
  - It shows 28 observations, of 3 variables, shoe_size, height, and gender. 
 
#b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.
```{r}
m_subset <- subset(data, gender == "M")
f_subset <- subset(data, gender == "F")

m_subset
f_subset
data
```
#c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
mean_of_shoe_size <- mean(shoe_size)
mean_of_height <- mean(height)
  
shoe_size_mean <- paste("Mean of shoe size",mean_of_shoe_size)
height_mean <- paste("Mean of height", mean_of_height)
  
shoe_size_mean
height_mean
```
#d. Is there a relationship between shoe size and height? Why?
#Yes, there is a relationship between shoe size and height because taller indiviaduals tend to have larger feet. 

#2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values. Consider data consisting of the names of months: ("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

```{r}
months <- c("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

factor_months <- factor(months)
factor_months
```
#3. Then check the summary() of the months_vector and factor_months_vector. Interpret the results of both vectors. Are they both equally useful in this case?

#The factor assigns levels to the months based on alphabetical order. A summary of the factor would show how many times each month appears in the vector.
```{r}
summary(months)
summary(factor_months)
```
#Create a vector and factor for the table below.
#Note: Apply the factor function with required order of the level. new_order_data <- factor(factor_data,levels = c("East","West","North")) print(new_order_data)
```{r}
direction <- c("East", "West", "North")
  frequency <- c(1, 4, 3)
  
data <- data.frame(direction, frequency)
data
  
new_order_data <- factor(direction,levels = c("East","West","North"))
print(new_order_data)
```
#5. Enter the data below in Excel with file name = import_march.csv
``` {r}
library(readxl)
excelData <- read.table("/cloud/project/Worksheet-4a/import_march.csv", header = TRUE, sep = ",")
```
#b. View the dataset. Write the R scripts and its result.
``` {r}
print(excelData)  
```

#6. 
#a. Create an R Program that allows the User to randomly select numbers from 1 to 50. Then display the chosen number. If the number is beyond the range of the selected choice,it will have to display a string “The number selected is beyond the range of 1 to 50”. If number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the number.
```{r}
input_num <- readline(prompt = "Enter a number between 1 and 50: ")
choose_num <- (1:50)
paste("Chosen number: ", choose_num)

if (input_num < 1 || input_num > 50) {
  print("The number selected is beyond the range of 1 to 50.")
} else if(input_num == 20) {
  print("True")
} else {
  input_num
}
```
#7. Change
#At ISATU University’s traditional cafeteria, snacks can only be purchased with bills. A long-standing rule at the concession stand is that snacks must be purchased with as few coins as possible. There are three types of bills: 50 pesos, 100 pesos, 200 pesos, 500 pesos, 1000 pesos.
#a. Write a function that prints the minimum number of bills that must be paid, given the price of the snack. Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills needed to purchase a snack.






