---
title: "Worksheet-4b in R"
author: "Trexan Sanico"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrix0 <- matrix (0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
  matrix0[i, j] <- abs(vectorA[i] - vectorA [j])
  }
}
matrix0
```
#2. Print the string "*" using for() function. The output should be the same as shown in Figure 1. 
```{r}
for (i in 1:5) {
  cat(rep('"*"',i), "\n")
}
```
#3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
a <- 0
b <- 1

c <- readline(prompt = "Enter a number: ")

repeat {
  c <- a + b
  if (c > 500) break
  a <- b
  b <- c
  print(c)
}
```
#4. Import the dataset as shown in Figure 1 you have created previously. 
#a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result


#b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.



