---
title: "Worksheet-2 in R"
author: "Trexan Sanico"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#1. Create a vector using : operator
#a. Sequence from -5 to 5. Write the R code and its output.
```{r}
num <- -5:5
num
```
#Describe its output.
#-The code displays a sequence from -5 to 5. 

#b. x <- 1:7. What will be the value of x?
```{r}
x <- 1:7
x
```
#-x will be a sequence of integers from 1 to 7.

#2. Create a vector using seq() function
#a. seq(1, 3, by=0.2) # specify step size. Write the R script and its output. 
```{r}
num <- seq(1, 3, by=0.2)
num
```
#Describe the output.
#-The output displays a sequence from 1 to 3 and was incremented by 0.2.

#3. A factory has a census of its workers. There are 50 workers in total. The following list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50,37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18.

#a. Access 3rd element, what is the value?
#-The value is 22. 
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50,37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
ages [3]
```

#b. Access 2nd and 4th element, what are the values?
-The vaues are 28 and 36.
```{r}
ages[c(2, 4)]
```

#c. Access all but the 4th and 12th element is not
included. Write the R script and its output.
```{r}
ages[-c(4, 12)]
```

#4. Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).
```{r}
x <- c("first" = 3, "second" = 0, "third" = 9)
print(x)
```

#a. Print the results. Then access x[c("first", "third")].
```{r}
names <- x[c("first", "third")]
names
```
#Describe the output.
#-The output shows a vector with named elements, and the values of the "first" and "third" elements. 

#5. Create a sequence x from -3:2.
#a. Modify 2nd element and change it to 0;
x[2] <- 0
x
```{r}
x <- -3:2
x[2] <- 0
x
```
#Describe the output.
#-The output displays a sequence from -3 to 2, and after the second
element was accessed, the -2 was changed to the value of 0. 

#6. 
#a. Create a data frame for month, price per liter (php)
and purchase-quantity (liter). Write the R scripts and
its output.
```{r}
month <- c("Jan", "Feb", "March", "Apr", "May", "June")
price <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
quant <- c(25, 30, 40, 50, 10, 45)
df <- data.frame(month, price, quant)
df
```

#b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use ‘weighted.mean(liter, purchase)‘. Write the R scripts and its output.
```{r}
weighted.mean(price, quant)
```

#7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.

#a. Type “rivers” in your R console. Create a vector data with 7 elements, containing the number of elements (length)in rivers, their sum (sum), mean (mean), median(median), variance(var), standard deviation(sd), minimum (min) and maximum (max).
```{r}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
data
```

#b. What are the results?
141.0000 (length of rivers)  
83357.0000 (sum of rivers)
591.1844 (mean of rivers)
425.0000 (median of rivers)
243908.4086 (variance of rivers)
493.8708 (standard deviation of rivers)
135.0000 (minimum length of rivers)
3710.0000 (maximum length of rivers)

#8. 
#a. Create vectors according to the above table.Write the R scripts and its output.
```{r}
celebrity <- c("Tom Cruise", "Rolling Stones","Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard  Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", "Bruce Springtsteen", "Dona;d Trump", "Muhammad Ali", "Paul McCarntney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K Rowling", "Brad Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant")
power_ranking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)
celeb_data <- data.frame(Celebrity = celebrity, Pay = pay, Power_ranking = power_ranking)
celeb_data
```

#b. Modify the power ranking and pay of J.K. Rowling. Change power ranking to 15 and pay to 90. Write the R scripts and its output.
```{r}
power_ranking[19] <- 15
pay[19] <- 90
celeb_data <- data.frame(Celebrity = celebrity, Pay = pay, Power_ranking = power_ranking)
celeb_data
```

#c. Create an excel file from the table above and save it
as csv file(PowerRanking). Import the csv file into
the RStudio. What is the R script?
```{r}
write.csv(celeb_data, file = "PowerRanking.csv", row.names = FALSE)
celeb_data_imported <- read.csv("PowerRanking.csv")
celeb_data_imported
```

#d. Access the rows 10 to 20 and save it as Ranks.RData. Write the R script and its output.
```{r}
ranks <- celeb_data[10:20,]
save(ranks, file = "Ranks.RData")
ranks
```

# e. Describe its output.
#The data contains rows 10 to 20, including the power rankings and pay of the celebrities.

#9.
#a. Import the excel file into your RStudio. What is the R script?
```{r}
library(readxl)
hotels_data <- read_excel("hotels-vienna.xlsx")
print(head(hotels_data))
```

#b. How many dimensions does the dataset have? What is the R script? What is its output?
```{r}
dimensions <- dim(hotels_data)
dimensions
```

#c. Select columns country, neighbourhood, price, stars, accomodation_type, and ratings. Write the R script.
```{r}
class(hotels_data)

selection <- hotels_data[c("country", "neighbourhood", "price", "stars", "accommodation_type", "rating")]
selection
```

#d. Save the data as *new.RData to your RStudio. Write the RScript. 
```{r}
save(selection, file = "new.RData")
```

#e. Display the first six rows and last six rows of the new.RData. What is the RScript?
```{r}
load("new.RData")
print(head(selection))
print(tail(selection))
```

#10. Create a list of ten (10) vegetables you ate during your lifetime. If none, just list down. 
#a. Write the R scripts and its output.
```{r}
veges <- c("squash", "carrots", "potato", "eggplant", "onion", "cabbage", "sayote", "moringa", "monggo", "stringbeans")
veges
```

#b. Add 2 additional vegetables after the last vegetables in the list. What is the Rscript and its output?
```{r}
veges <- c(veges, "okra", "pechay")
veges
```

#c Add 4 additional vegetables after index 5. How many data points does your vegetable list have? What is the R script and its output?
```{r}
veges <- append(veges, c("onions", "cucumber", "mushroom", "radish"), after = 5)
veges
```

#d. Remove the vegetables in index 5, 10, and 15. How many vegetables were left? Write the codes and its output.
```{r}
veges <- veges[-c(5, 10, 15)]
veges
```