---
title: "Worksheet-1 in R"
author: "Trexan Sanico"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. 
a. How many data points? 
+There are 34 data points.

b. Write the R code and its output.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)

points <- length(age)

print(points)
```

2. Find the reciprocal of the values for age.
```{r}
reciprocal_age <- 1 / age

print(reciprocal_age)
```

3. What happened to the new_age?
+It duplicates the age vector with 0 in the middle. 

```{r}
new_age <- c(age, 0, age) 

print(new_age)
```

4. Sort the values for age
```{r}
sortAge <- sort(age)

print(sortAge)
```

5. Find the minimum and maximum value for age. 
```{r}
minAge <- min(age)
minAge

maxAge <- max(age)
maxAge
```

6.
a. How many data points? 
+There are 12 data points.

b. Write the R code and its output.
```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)

points <- length(data)

print(points)
```

7.  Generate a new vector for data where you double every value of the data. What happen to the data?
+The data were increased by two.
```{r}
double_data <- data * 2

print(double_data)
```

8. Generate a sequence for the following scenario:
8.1 Integers from 1 to 100.
```{r}
sequence <- seq(1, 100)

print(sequence)
```

8.2 Numbers from 20 to 60
```{r}
n <- seq(20, 60)

print(n)
```

8.3 Mean of numbers from 20 to 60
```{r}
numSeq <- seq(20, 60)

meanNum <- mean(numSeq)
print(meanNum)
```

8.4 Sum of numbers from 51 to 91
```{r}
numsum <- sum(seq(51, 91))
numsum
```

8.5 Integers from 1 to 1,000
```{r}
s <- seq(1:1000)

print(s)
```

a. How many data points from 8.1 to 8.4?
+There are 143 data points. 
```{r}
length1 <- length(sequence)
length2 <- length(n)
length3 <- length(meanNum)
length4 <- length(numsum)
sum (length1+length2+length3+length4)
```

b. Write the R code and its output from 8.1 to 8.4.
```{r}
sequence <- seq(1:100)
n <- seq(20:60)
meanNum <- mean(20:60)
numsum <- sum(51:91)
```

c. For 8.5 find only maximum data points until 10.
```{r}
m <- 1:1000
ans <- max(m[t <- 10])
ans
```

9. Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and 7 using filter option.
```{r}
Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
```

10. Generate a sequence backwards of the integers from 1 to 100.
```{r}
bckwrd <- seq(100, 1, by = -1)

print(bckwrd)
```

11. List all the natural numbers below 25 that are multiples of 3 or 5. Find the sum of these multiples.
```{r}
numbers <- 1:24

multiples <- numbers[numbers %% 3 == 0 | numbers %% 5 == 0]
print(multiples)

sum <- sum(multiples)
print(sum)
```

a. How many data points from 10 to 11?
+There are 12 data points.

b. Write the R code and its output from 10 and 11.
```{r}
dp10 <- length(multiples)
dp11 <- 1

total <- dp10 + dp11
total
```


13. Find x[2] and x[3]. Write the R code and its output. 
```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75,
75, 77)

x2 <- score[2]
x3 <- score[3]

print(x2)
print(x3)
```

14.Create a vector a = c(1,2,NA,4,NA,6,7).
a. Change the NA to 999 using the codes print(a,na.print="-999").
b. Write the R code and its output. Describe the output.
+The output displays the vector 'a' with NA values being replaced by '-999'. The values of a in memory remain the same. 
```{r}
a <- c(1,2,NA,4,NA,6,7)

print(a, na.print = "-999")
```

15. 
```{r}
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
```
What is the output of the above code?
+I input my name and my age and it shows as this, "My name is trexan sanico and I am 19 years old."


